---
title: "IRIS"
author: "Pramod Verma"
date: "16 March 2018"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is the R Markdown file created for introducing the basic building blocks of R to PSA Data Analytics Technical Participants.

```{r message=FALSE, warning=FALSE}
```

```{r iris}

head(iris)
str(iris)
dim(iris)
names(iris)
class(iris)
summary(iris)
```

## Assigning Variables and Plotting

```{r }

x <- iris$Sepal.Length
y <- iris$Sepal.Width
z <- iris$Species

plot(x)
hist(x, col = "red")
plot(x,y)
boxplot(x)

```

## Looking at all the species in the same scatterplot for Petal Length & Width

```{r }
plot(iris$Petal.Length, iris$Petal.Width, pch=21, bg=c("red","green3","blue")[unclass(iris$Species)], 
main="Edgar Anderson's Iris Data")
```


## Looking at the scatterplot of all the variables for each species

```{r }
pairs(iris[1:4], main = "Edgar Anderson's Iris Data", pch = 21, 
bg = c("red", "green3", "blue")[unclass(iris$Species)])

```

## Creating pi chart to show the share of each species

```{r }
pie(table(iris$Species), 
main = "Pie Chart of the Iris data set Species", 
col = c("orange1", "chocolate", "coral"), 
radius = 1)

```

## Using for loop to create multiple plots

```{r }
par(mfrow=c(1,4)) 
color <- c("red", "blue","green","brown") 
for(i in 1:4) { 
hist(iris[,i], main=names(iris)[i], col = color[i]) }

```

## Finding out the correlation among all the numerical variables

```{r message=FALSE, warning=FALSE}
cor(iris[,c(1:4)])
library(PerformanceAnalytics)
chart.Correlation(iris[,c(1,2,3,4)], histogram=TRUE, pch=19)

```

## Demostration of K Means Clustering

```{r }
normalize <- function(x){
  return ((x-min(x))/(max(x)-min(x)))
}

iris$Sepal.Length<- normalize(iris$Sepal.Length)
iris$Sepal.Width<- normalize(iris$Sepal.Width)
iris$Petal.Length<- normalize(iris$Petal.Length)
iris$Petal.Width<- normalize(iris$Petal.Width)
head(iris)

result<- kmeans(iris[,c(1,2,3,4)],3)


plot(iris[c(1,2)], col=result$cluster, pch = 19, main =" Clusters based on Sepal Length and Width")

plot(iris[c(1,2)], col=iris$Species, pch = 19, main =" Clusters based on Flower Species")

plot(iris[c(3,4)], col=result$cluster, pch = 19, main =" Clusters based on Petal Length and Width")

plot(iris[c(3,4)], col=iris$Species, pch = 19, main =" Clusters based on Flower Species")
```

## Running the k means algorithms four times on the same dataset

```{r }
for (i in 1:4) {
  set.seed(100 +50*i)
  result<- kmeans(iris[,c(1,2,3,4)],3)
plot(iris[c(3,4)], col=result$cluster, pch = 19, main =c(" Clusters based on Petal Length and Width: Iteration ", i)) }

```
#### It shows the importance of multiple iteration and averaging them

## End of the Script